---
title: "Exponential Distribution vs. Central Limit Theorem"
author: "Richard Nacianceno"
date: "5/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Environment

```{r Envir, message=FALSE}
library(tidyverse)
library(glue)
library(matrixStats)
downloadDate <- date()

glue("Simulation date is {downloadDate}")
sessionInfo()
```

### About

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations. 

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

1. Show the sample mean and compare it to the theoretical mean of the distribution.

2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

3. Show that the distribution is approximately normal.

## Simulation

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials.

```{r SimulationDatasets}
set.seed(420) # random number generator set at 420
lambda <- .2 
sampCol <- 40 # number of variables
sampRow <- 1000  # number of observations

expoDis <- matrix(rexp(n=sampRow*sampCol, rate = lambda), sampRow, sampCol)

expoDF <- as.tibble(expoDis)
# Converted it to a tibble to make it easier to visualize

e_df <- expoDF %>% mutate("MEAN" = rowMeans(expoDis), "SD" = rowSds(expoDis), "SumCumu" = cumsum(MEAN)/(1:sampRow)) %>% select(MEAN, SD, SumCumu)
```

## 1. Sample mean vs. the theoretical mean of the distribution.

```{r E_mean vs T_mean}
e_mean <- mean(e_df$MEAN)
t_mean <- 1/lambda
res_df <- tibble(Result = c("Sample means", "Theoretical mean"),"Mean"=c(e_mean, t_mean))
res_df
```
As we can see above, the exponential distribution mean is `r e_mean` is close to the CLT mean of `r t_mean`. The graph below shows the cumulative sum of the sample mean approaching the horizontal line valuing at 1/lambda which is our theoritcal mean.  

```{r expo vs Theo}
ggplot(data=e_df, aes(x = 1:sampRow, y = SumCumu))+ g1 + geom_hline(yintercept = t_mean) +
geom_line(color="coral", size = 3) + labs(x = "Number of obs", y = "Cumulative mean", title = "Sample Means of 1000 Observation and 40 Variables")


```
Notes:

1. Using (Freedman-Diaconis)[https://en.wikipedia.org/wiki/Freedman%E2%80%93Diaconis_rule] rule to select the width of the bins inside a histogram.  

```{r error=TRUE}
binhist <- hist(expoDF$MEAN, plot=FALSE, breaks = "FD")
g2 <- ggplot(data = expoDF, aes(x=MEAN,y=..density..)) 
g2 + geom_histogram(fill="white", col="black",breaks =binhist$breaks)+geom_density(aes(MEAN), color = "blue")+geom_vline(xintercept = c(e_mean, t_mean), color = c("red", "purple")) + labs(x= "Sample Means", y= "Density", title = "Histogram of Sample Means")
```




